---
title: "Introduction to the rtZIKVrisk R package"
author: "Spencer J. Fox"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the rtZIKVrisk R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Background
rtZIKVrisk was created to accompany the results from our study [Real-time Zika Risk Assessment in the United States](http://biorxiv.org/content/early/2016/06/07/056648). It's meant to allow users to recreate the figures from the manuscript, and also allows for one to simulate zika epidemics with our framework and analyze the resultant outbreaks.

## Installation (R)
To install the package use the [devtools](https://cran.r-project.org/web/packages/devtools/index.html) package on CRAN.
```r
    install.packages("devtools")
    
    devtools::install_github("sjfox/rtZIKVrisk")
```

Once the package has been installed, it can be loaded with 

```{r, warning=FALSE, message=FALSE, comment=FALSE}
    library(rtZIKVrisk)
```

## Simulating Zika outbreaks
The package allows one to simulate outbreaks according to the Zika outbreak simulation model presented in our paper, according to default parameters or other combinations. Normally we use the `run_n_zika_sims()` function to run multiple outbreaks, so let's run 10 stochastic realizations with the default parameter values, and plot the histogram of final sizes, and the number of reported autochthonous (locally transmitted) cases.

```{r, fig.width=6}
  set.seed(808)
  sim_parms <- zika_def_parms()
  outbreak_sim <- run_n_zika_sims(num_reps = 10, sim_parms)
  plot_final_sizes(outbreak_sim)
  plot_zika_outbreaks(outbreak_sim)
  
```
